---
title: "CL and COMLAND Extractions"
author: "Alex Hanke"
date: "April 20, 2016"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(RODBC)
channel <- odbcConnect("ptran", uid="hankea", pwd="StAndr3ws",believeNRows=F)
```

## COMLAND
The commercial data prior to 2002 resides in COMLAND. MARFIS is the current database. These commercial data sources have a 2 year overlap, roughly 2001 and 2002.

```{r}
sqlTables(channel,schema="COMLAND")
sqlTables(channel,schema="CL")
sqlTables(channel,schema="HANKEA")
```

It appears that a separate table is created for each year. The sub_trips tables describe the fishing position, gear, timing and effort.
```{r}
sqlQuery(channel,"select * from COMLAND.SUB_TRIPS_1999 where rownum<5 and gear_type=51")
sqlQuery(channel,"select * from CL.SUB_TRIPS_1999 where rownum<5 and gear_type=51")

sqlQuery(channel,"select * from CL.ZIFDB19862008 where rownum < 6 and GEARCODE = 51")
sqlQuery(channel,"select * from CL.ZIFNHQ1999 where rownum < 6 and GEARCODE = 51")

```

The identified_catches provides set wise information on the weight of each species caught and landing time and place.
```{r}
sqlQuery(channel,"select * from COMLAND.IDENTIFIED_CATCHES_1999 where rownum<5 ")
sqlQuery(channel,"select * from CL.IDENTIFIED_CATCHES_1999 where rownum<5 ")
```

The cl_log_header contains information on the gear.
```{r}
sqlQuery(channel,"select * from COMLAND.CL_LOG_HEADER_1995 where rownum<5 ")
sqlQuery(channel,"select * from COMLAND.CL_LOG_DETAIL_1995 where rownum<5 ") # does not exist!

sqlQuery(channel,"select * from CL.LOG_HEADER_1999 where rownum<5 ")
sqlQuery(channel,"select * from CL.LOG_DETAIL_1999 where T_NBR_HOOKS>10 and rownum<100 ")
sqlQuery(channel,"select * from CL.LOG_ESTIMATE_1999 where rownum<5 ")

sqlQuery(channel,"select * from CL.SLIP_HEADER_1999 where rownum<5 ")
sqlQuery(channel,"select * from CL.SLIP_DETAIL_1999 where rownum<5 ")
```

The catchers tables provide vessel info.
```{r}
sqlQuery(channel,"select * from COMLAND.CATCHERS_1999 where rownum<5 ")
sqlQuery(channel,"select * from CL.CATCHERS_1999 where rownum<5 ")
```

The other tables describe the codes.
```{r}
sqlQuery(channel,"select * from COMLAND.NAFO_CODES")
sqlQuery(channel,"select * from COMLAND.NAFO_GEAR_CODES")
sqlQuery(channel,"select * from COMLAND.NAFO_SPECIES_CODES")
sqlQuery(channel,"select * from COMLAND.NAFO_MAIN_SPECIES_CODES")
sqlQuery(channel,"select * from COMLAND.SPECIES_PRE_1986")
sqlQuery(channel,"select * from COMLAND.GEAR_TYPES_PRE_1986")
sqlQuery(channel,"select * from COMLAND.SPECIES_CODE_LINK")

sqlQuery(channel,"select * from CL.STATISTICAL_DISTRICTS")
sqlQuery(channel,"select * from CL.NAFO_DIVISIONS")
sqlQuery(channel,"select * from CL.NAFO_SUB_AREAS")
sqlQuery(channel,"select * from CL.NAFO_UNIT_AREAS")
sqlQuery(channel,"select * from CL.DFO_REGIONS")
sqlQuery(channel,"select * from CL.COMMUNITIES where rownum < 6")
sqlQuery(channel,"select * from CL.FORMS")


sqlQuery(channel,"select * from CL.NAFO_SPECIES_CODES")
sqlQuery(channel,"select * from CL.NAFO_MAIN_SPECIES_CODES")
sqlQuery(channel,"select * from CL.SPECIES_PRE_1986")
sqlQuery(channel,"select * from CL.GEAR_TYPES_PRE_1986")
sqlQuery(channel,"select * from CL.GEAR_TYPES")
sqlQuery(channel,"select * from CL.SPECIES")
sqlQuery(channel,"select * from CL.SIZES")

sqlQuery(channel,"select SPECIES_CODE, DESC_ENG from MARFISSCI.SPECIES order by SPECIES_CODE asc")
```
41           251                  2                      Swordfish   AAA
42           252                  2                  Albacore tuna   AAA
43           253                  2                    Bigeye tuna   AAA
44           254                  2                   Bluefin tuna   AAA
45           255                  2                 Tuna, skipjack   AAA
46           256                  2                 Yellowfin tuna   AAA
47           259                  2               Unspecified tuna   AAA
48           260                  2                       Billfish   AAA
49           261                  2                    Blue marlin   AAA
50           262                  2                   White marlin   AAA
51           299                  2            Unspecified pelagic   AAA



Let's try to grab some swordfish data.
23   49    LL*                            LONGLINES (CHARTERS)
24   50     LL                        LONGLINE (NOT SPECIFIED)
25   51    LLS     SET LINES (BOTTOM OR NEAR BOTTOM LONGLINES)
26   52    LLD                DRIFT LINES (DRIFTING LONGLINES)

sqlQuery(channel,"select distinct(gear_id) from CL.LOG_DETAIL_1999 where T_NBR_HOOKS>100 ")
  GEAR_ID
1    5100
2    5900
3    6000
4    8100
```{r}
SWO = sqlQuery(channel,"select * from comland.S_1986_2001 where gear_type in (49,50,51,52)")
SPECIES = sqlQuery(channel,"select SPECIES_CODE, DESC_ENG from MARFISSCI.SPECIES order by SPECIES_CODE asc")
SWO = merge(SWO,SPECIES,by.x="MAIN_SPECIES_CAUGHT",by.y="SPECIES_CODE",all.x=T)
xtabs(~DESC_ENG+YEAR_OF_ACTIVITY,SWO)
```
or

Note: It seems LOG_CODE and TRIP_ID group the data in the same way. ESTIMATE_KEY seems to describe sets within the trips and this key is what you use to link to other tables.
```{r}
SWO = as.data.table(sqlQuery(channel,"select * from CL.LOG_DETAIL_1999 where T_NBR_HOOKS>10"))
a = SWO[GEAR_ID==5100,.N,by=.(TRIP_CODE,LOG_CODE)]
SWO[TRIP_CODE==a[83,TRIP_CODE],] # one trip with 5 rows 
b = SWO[LOG_CODE==a[83,LOG_CODE],ESTIMATE_KEY]

sqlQuery(channel,paste("select * from CL.LOG_DETAIL_1999 where estimate_key in ('",paste(b,collapse="','"),"') ",sep="" ))
sqlQuery(channel,paste("select * from CL.LOG_ESTIMATE_1999 where estimate_key in ('",paste(b,collapse="','"),"') ",sep="" ))


SWO = sqlQuery(channel,"select * from cl.ALL_SUB_TRIPS_VIEW where gear_type in (49,50,51,52)")
SPECIES = sqlQuery(channel,"select SPECIES_CODE, DESC_ENG from MARFISSCI.SPECIES order by SPECIES_CODE asc")
SWO = merge(SWO,SPECIES,by.x="MAIN_SPECIES_CAUGHT",by.y="SPECIES_CODE",all.x=T)
xtabs(~DESC_ENG+YEAR_OF_ACTIVITY,SWO)
```

# This portion of the script was given to me by Mark Fowler.
```{r cars}
#create annual tables of log-based catch-effort data (with some help from the slip data)
#This job works for 1986-1995 and 1998-2002 (1996-1997 were missed when formatting was amended to deal with 4-digit years)
#This job is run in SQL-Plus via '@d:\assess\extract\marfis\ziff\testrun9697.sql'
Year = 1999
sqlQuery(channel,"drop table agglog")
sqlQuery(channel, paste("create table agglog as select d.log_code lcode, ",
  "(substr(d.log_code,1,12) || substr(d.log_code,14,5)) logcode, ",
  "substr(e.t_species_id_est,1,3) spec, ",
  "sum(e.t_live_weight/1000) logwt ",
  "from cl.log_detail_",Year," d, ",
  "cl.log_estimate_",Year," e ",
  "where d.log_code=e.log_code ",
  "and d.estimate_key=e.estimate_key ",
  "group by d.log_code, substr(e.t_species_id_est,1,3)", sep=""))

#Ignoring gear on the slip for selecting, as it is often incorrect.
sqlQuery(channel,"drop table slip")
sqlQuery(channel, paste("create table slip as select (substr(slip_code,1,12) || substr(slip_code,14,5)) slipcode, ",
  "gear_id gear,substr(species_id,1,3) spec, sum(t_live_weight/1000) slipwt ",
  "from cl.slip_detail_",Year," ",
  "group by (substr(slip_code,1,12) || substr(slip_code,14,5)), ",
  "gear_id,substr(species_id,1,3)", sep=""))

#Both slip and log denote the gear, making it seem a likely link for joining. However
#the gear type is miscoded on the slip with extreme frequency, as evidenced by the
#nature of the effort data in the log (typically suitable for the log gear type when
#log and slip disagree). Thus gear is discounted when joining.
#and l.gear=s.gear (+)
#Have to edit for desired species codes. The structure facilitates keeping track of species composition in the catch. The TotalWt is for all species, for whatever aggregation the source data represents (set, trip, etc), so can be used to check if a given record might be shy a relevant species. I.e. it is not the summed weight of the species selected, but of all species in the catch.
#Originally written for CPUE analyses, with following filters.
#and gear_id in ('1200','1500','1900','2100','3100','4100','5100','6200','6300','7100')
#and t_live_weight>0
#and (t_nbr_hours>0 or t_nbr_sets>0 or t_nbr_tows>0 or t_nbr_hooks>0)

sqlQuery(channel,"drop table logadjslip")
sqlQuery(channel, paste("create table logadjslip as",
  "select l.lcode,l.logcode,l.spec,logwt,",
  "nvl(slipwt,0) slipwt, nvl((nvl(slipwt,0)/nullif(logwt,0)),1) adjfac",
  "from agglog l, slip s",
  "where logcode=slipcode (+)",
  "and l.spec=s.spec (+)"))

sqlQuery(channel,"drop table cpuelog")
sqlQuery(channel, paste("create table cpuelog as ",
  "select h.log_code logcode, e.estimate_key ekey, substr(h.log_code,1,6) cfv, ",
  "unit_area area,  date_fished, t_latitude lat, ",
  "t_longitude lon, substr(gear_id,1,2) gear,(t_depth/1.83) depth, ",
  "effort_measure eff, t_nbr_hours hrs, t_nbr_sets sets, ",
  "t_nbr_tows tows,t_nbr_hooks hooks, unit_of_measure units, ",
  "nvl(sum(nvl((t_live_weight/1000),0)),0) TotalWt, ",
  "nvl(sum(decode(substr(t_species_id_est,1,3),'351', nvl((t_live_weight/1000), 0))),0) Argentine, ",
  "nvl(sum(decode(substr(t_species_id_est,1,3),'360', nvl((t_live_weight/1000), 0))),0) Capelin, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'100', nvl((t_live_weight/1000), 0))),0) Cod, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'173', nvl((t_live_weight/1000), 0))),0) Cusk, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'362', nvl((t_live_weight/1000), 0))),0) Dogfish, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'143', nvl((t_live_weight/1000), 0))),0) WinterF, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'142', nvl((t_live_weight/1000), 0))),0) WitchF, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'141', nvl((t_live_weight/1000), 0))),0) YellowtailF, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'179', nvl((t_live_weight/1000), 0))),0) Grenadier, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'110', nvl((t_live_weight/1000), 0))),0) Haddock, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'172', nvl((t_live_weight/1000), 0))),0) HakeS, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'171', nvl((t_live_weight/1000), 0))),0) HakeW, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'130', nvl((t_live_weight/1000), 0))),0) Halibut, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'200', nvl((t_live_weight/1000), 0))),0) Herring, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'177', nvl((t_live_weight/1000), 0))),0) Monkfish, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'140', nvl((t_live_weight/1000), 0))),0) Plaice, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'170', nvl((t_live_weight/1000), 0))),0) Pollock, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'120', nvl((t_live_weight/1000), 0))),0) Redfish, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'372', nvl((t_live_weight/1000), 0))),0) SharkB, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'369', nvl((t_live_weight/1000), 0))),0) SharkP, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'356', nvl((t_live_weight/1000), 0))),0) Skate, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'144', nvl((t_live_weight/1000), 0))),0) Turbot, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'174', nvl((t_live_weight/1000), 0))),0) Wolffish, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'250', nvl((t_live_weight/1000), 0))),0) Mackerel, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'608', nvl((t_live_weight/1000), 0))),0) SurfClam, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'612', nvl((t_live_weight/1000), 0))),0) Scallop, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'613', nvl((t_live_weight/1000), 0))),0) Squid, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'617', nvl((t_live_weight/1000), 0))),0) Whelk, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'700', nvl((t_live_weight/1000), 0))),0) Lobster, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'702', nvl((t_live_weight/1000), 0))),0) Shrimp, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'705', nvl((t_live_weight/1000), 0))),0) SnowCrab, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'703', nvl((t_live_weight/1000), 0))),0) JonahCrab, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'712', nvl((t_live_weight/1000), 0))),0) PMont, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'251', nvl((t_live_weight/1000), 0))),0) Swordfish, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'252', nvl((t_live_weight/1000), 0))),0) TunaAlb, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'253', nvl((t_live_weight/1000), 0))),0) TunaBE, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'254', nvl((t_live_weight/1000), 0))),0) TunaBF, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'255', nvl((t_live_weight/1000), 0))),0) TunaSJ, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'256', nvl((t_live_weight/1000), 0))),0) TunaYF, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'259', nvl((t_live_weight/1000), 0))),0) TunaUnk, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'375', nvl((t_live_weight/1000), 0))),0) SharkM, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'379', nvl((t_live_weight/1000), 0))),0) SharkUnk, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'350', nvl((t_live_weight/1000), 0))),0) Alewife,",
"nvl(sum(decode(substr(t_species_id_est,1,3),'149', nvl((t_live_weight/1000), 0))),0) UNKFlounder ",
  "from cl.log_header_",Year," h,cl.log_detail_",Year," d,cl.log_estimate_",Year," e ",
  "where h.log_code=d.log_code ",
  "and h.log_code=e.log_code ",
  "and d.estimate_key=e.estimate_key ",
  "group by e.estimate_key, h.log_code, unit_area,date_fished, ",
  "t_latitude, t_longitude,gear_id,(t_depth/1.83), ",
  "effort_measure,t_nbr_hours,t_nbr_sets,t_nbr_tows, ",
  "t_nbr_hooks,t_lines_per_tub, t_hooks_per_line,unit_of_measure",sep=""))

sqlQuery(channel, "drop table cpueslip")
sqlQuery(channel, paste("create table cpueslip as select h.log_code logcode, ",
  "e.estimate_key ekey, substr(h.log_code,1,6) cfv,unit_area area,  date_fished, ",
  "t_latitude lat, t_longitude lon,substr(gear_id,1,2) gear,(t_depth/1.83) depth, ",
  "effort_measure eff,t_nbr_hours hrs,t_nbr_sets sets,t_nbr_tows tows, ",
  "t_nbr_hooks hooks,unit_of_measure units, ",
  "nvl(sum(nvl((t_live_weight/1000),0)),0) TotalWt, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'351', nvl((adjfac*t_live_weight/1000), 0))),0) Argentine, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'360', nvl((adjfac*t_live_weight/1000), 0))),0) Capelin, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'100', nvl((adjfac*t_live_weight/1000), 0))),0) Cod, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'173', nvl((adjfac*t_live_weight/1000), 0))),0) Cusk, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'362', nvl((adjfac*t_live_weight/1000), 0))),0) Dogfish, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'143', nvl((adjfac*t_live_weight/1000), 0))),0) WinterF, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'142', nvl((adjfac*t_live_weight/1000), 0))),0) WitchF, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'141', nvl((adjfac*t_live_weight/1000), 0))),0) YellowtailF, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'179', nvl((adjfac*t_live_weight/1000), 0))),0) Grenadier, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'110', nvl((adjfac*t_live_weight/1000), 0))),0) Haddock, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'172', nvl((adjfac*t_live_weight/1000), 0))),0) HakeS, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'171', nvl((adjfac*t_live_weight/1000), 0))),0) HakeW, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'130', nvl((adjfac*t_live_weight/1000), 0))),0) Halibut, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'200', nvl((adjfac*t_live_weight/1000), 0))),0) Herring, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'177', nvl((adjfac*t_live_weight/1000), 0))),0) Monkfish, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'140', nvl((adjfac*t_live_weight/1000), 0))),0) Plaice, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'170', nvl((adjfac*t_live_weight/1000), 0))),0) Pollock, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'120', nvl((adjfac*t_live_weight/1000), 0))),0) Redfish, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'372', nvl((adjfac*t_live_weight/1000), 0))),0) SharkB, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'369', nvl((adjfac*t_live_weight/1000), 0))),0) SharkP, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'356', nvl((adjfac*t_live_weight/1000), 0))),0) Skate, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'144', nvl((adjfac*t_live_weight/1000), 0))),0) Turbot, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'174', nvl((adjfac*t_live_weight/1000), 0))),0) Wolffish, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'250', nvl((adjfac*t_live_weight/1000), 0))),0) Mackerel, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'608', nvl((adjfac*t_live_weight/1000), 0))),0) SurfClam, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'612', nvl((adjfac*t_live_weight/1000), 0))),0) Scallop, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'613', nvl((adjfac*t_live_weight/1000), 0))),0) Squid, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'617', nvl((adjfac*t_live_weight/1000), 0))),0) Whelk, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'700', nvl((adjfac*t_live_weight/1000), 0))),0) Lobster, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'702', nvl((adjfac*t_live_weight/1000), 0))),0) Shrimp, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'705', nvl((adjfac*t_live_weight/1000), 0))),0) SnowCrab, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'703', nvl((adjfac*t_live_weight/1000), 0))),0) JonahCrab, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'712', nvl((adjfac*t_live_weight/1000), 0))),0) PMont, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'251', nvl((adjfac*t_live_weight/1000), 0))),0) Swordfish, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'252', nvl((adjfac*t_live_weight/1000), 0))),0) TunaAlb, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'253', nvl((adjfac*t_live_weight/1000), 0))),0) TunaBE, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'254', nvl((adjfac*t_live_weight/1000), 0))),0) TunaBF, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'255', nvl((adjfac*t_live_weight/1000), 0))),0) TunaSJ, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'256', nvl((adjfac*t_live_weight/1000), 0))),0) TunaYF, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'259', nvl((adjfac*t_live_weight/1000), 0))),0) TunaUnk, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'375', nvl((adjfac*t_live_weight/1000), 0))),0) SharkM, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'379', nvl((adjfac*t_live_weight/1000), 0))),0) SharkUnk, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'350', nvl((adjfac*t_live_weight/1000), 0))),0) Alewife, ",
"nvl(sum(decode(substr(t_species_id_est,1,3),'149', nvl((adjfac*t_live_weight/1000), 0))),0) UNKFlounder ",
  "from cl.log_header_",Year," h,cl.log_detail_",Year," d,cl.log_estimate_",Year," e, logadjslip l ",
  "where h.log_code=d.log_code " ,
  "and h.log_code=e.log_code ",
  "and h.log_code=l.lcode ",
  "and d.estimate_key=e.estimate_key ",
  "and substr(t_species_id_est,1,3)=spec (+)",
  "group by e.estimate_key, h.log_code, unit_area,date_fished, ",
  "t_latitude, t_longitude,gear_id,(t_depth/1.83), ",
  "effort_measure,t_nbr_hours,t_nbr_sets,t_nbr_tows, ",
  "t_nbr_hooks,t_lines_per_tub, t_hooks_per_line,unit_of_measure",sep=""))

#Based on 2000, the effort variable possibilities sets, tows and hooks
# were mutually exclusive. Only one ever existed, hours being the sole
# likelihood of a second effort variable. This fits with the ZIF practice
# of having just the two hours and count variables. And note that the count
# variable will represent sheets for gillnets, hooks for longliners.
#Remember that the t_live_weight values are in kgs. Do not let the existence
# of the units_of_measure variable confuse, it relates to the skippers estimates.
#Note bogus data for catchers_recid, trip_sub_trip_flag, trip_num, sub_trip_num. [For a VDC application]
# Subtrips and fishing days will be computed by the application as and when requested. [This comment refers to a VDC application]
#Tossed completely zero catch records for CPUE analyses with code below
#and (Argentine+Capelin+Cod+Cusk+Dogfish+WinterF+WitchF+YellowtailF+
#Grenadier+Haddock+HakeS+HakeW+Halibut+Herring+Monkfish+Plaice+Pollock+Redfish+SharkB+
#SharkP+Skate+Turbot+Wolffish+Mackerel+SurfClam+Scallop+Squid+Whelk+Lobster+Shrimp+SnowCrab+PMont+Swordfish+Alewife)>0;

sqlQuery(channel,"drop table cpueloglong")
sqlQuery(channel, paste("create table cpueloglong as select logcode, ekey,  ",
  "'XXXXXXXXXXXX' catchers_recid,'S' region_code, cfv cfv_number, ",
  "to_number('19' || substr(logcode,9,2)) yland, ", 
  "to_number(substr(logcode,11,2)) mland, ",
  "to_number(substr(logcode,13,2)) dland, ",
  "date_fished caught_date, substr(area,1,3) nafo_division_code, ",
  "lower(substr(area,4,1)) nafo_unit_area, tonnage_class_code tonnage_class, ",
  "length_class_code length_class, gear gear_type, ",
  "decode(greatest(Argentine,Capelin,Cod,Cusk,Dogfish,WinterF,WitchF,YellowtailF,Grenadier,Haddock,HakeS,HakeW,Halibut,Herring,Monkfish,Plaice,Pollock,Redfish,SharkB,SharkP,Skate,Turbot,Wolffish,Mackerel,SurfClam,Scallop,Squid,Whelk,Lobster,Shrimp,SnowCrab,JonahCrab,PMont,Swordfish,TunaBE,TunaBF,SharkM,Alewife,UNKFlounder,TunaAlb,TunaSJ,TunaYF,TunaUnk,SharkUnk),Argentine,'351',Cod,'100',Capelin,'360',Cusk,'173',Dogfish,'362',WinterF,'143',WitchF,'142',YellowtailF,'141',Grenadier,'179',Haddock,'110',HakeS,'172',HakeW,'171',Halibut,'130',Herring,'200',Monkfish,'177',Plaice,'140',Pollock,'170',Redfish,'120',SharkB,'372',SharkP,'369',Skate,'356',Turbot,'144',Wolffish,'174',Mackerel,'250',Swordfish,'251',TunaBE,'253',TunaBF,'254',SharkM,'375',Alewife,'350',SurfClam,'608',Scallop,'612',Squid,'613',Whelk,'617',Lobster,'700',Shrimp,'702',SnowCrab,'705',JonahCrab,'703',PMont,'712',UNKFlounder,'149',TunaAlb,'252',TunaSJ,'255',TunaYF,'256',TunaUnk,'259',SharkUnk,'379','000') main_species_caught," ,
  "to_number(substr(to_char(date_fished,'YYYY'),1,4)) ycaught, ",
  "to_number(substr(to_char(date_fished,'MM'),1,2)) mcaught, ",
  "to_number(substr(to_char(date_fished,'DD'),1,2)) dcaught, ",
  "0 trip_num,0 sub_trip_num,depth,decode((-1), ",
  "sign(depth-.01),'10', sign((depth-.01)*(depth-25.99)),'1', ",
  "sign((depth-25.995)*(depth-50.99)),'2', sign((depth-50.995)*(depth-75.99)),'3', ",
  "sign((depth-75.995)*(depth-100.99)),'4',sign((depth-100.995)*(depth-125.99)),'5', ",
  "sign((depth-125.995)*(depth-150.99)),'6', ",
  "sign((depth-150.995)*(depth-175.99)),'7', ",
  "sign((depth-175.995)*(depth-200.99)),'8', ",
  "sign((depth-200.995)*(depth-250.99)),'9', ",
  "sign((depth-250.995)*(depth-997.99)),'0','10') depthzone, ",
  "lat latitude,lon longitude, 0 fish_days, hrs effort_hrs, ",
  "greatest(sets, tows, hooks) effort_count,'XX' trip_sub_trip_flag, ",
  "TotalWt,Argentine,Capelin,Cod,Cusk,Dogfish,WinterF,WitchF,YellowtailF,
Grenadier,Haddock,HakeS,HakeW,Halibut,Herring,Monkfish,Plaice,Pollock,Redfish,SharkB,
SharkP,Skate,Turbot,Wolffish,Mackerel,SurfClam,Scallop,Squid,Whelk,Lobster,Shrimp,SnowCrab,JonahCrab,PMont,Swordfish,TunaBE,TunaBF,SharkM,Alewife,UNKFlounder,TunaAlb,TunaSJ,TunaYF,TunaUnk,SharkUnk ",
  "from cpuelog, cl.catchers_",Year, "  ",
  "where cfv = cfv_number (+) ", sep=""))

#filter for CPUE analyses
#and (Argentine+Capelin+Cod+Cusk+Dogfish+WinterF+WitchF+YellowtailF+
#Grenadier+Haddock+HakeS+HakeW+Halibut+Herring+Monkfish+Plaice+Pollock+Redfish+SharkB+
#SharkP+Skate+Turbot+Wolffish+Mackerel+SurfClam+Scallop+Squid+Whelk+Lobster+Shrimp+SnowCrab+PMont+Swordfish+Alewife)>0;

sqlQuery(channel,"drop table cpuesliplong") 
sqlQuery(channel, paste("create table cpuesliplong as select logcode,ekey, ",
  "'XXXXXXXXXXXX' catchers_recid,'S' region_code, cfv cfv_number, ",
  "to_number('19' || substr(logcode,9,2)) yland, ",
  "to_number(substr(logcode,11,2)) mland, ",
  "to_number(substr(logcode,13,2)) dland, ",
  "date_fished caught_date, substr(area,1,3) nafo_division_code, ",
  "lower(substr(area,4,1)) nafo_unit_area, nvl(tonnage_class_code,0) tonnage_class, ",
  "nvl(length_class_code,0) length_class, gear gear_type, ",
  "decode(greatest(Argentine,Capelin,Cod,Cusk,Dogfish,WinterF,WitchF,YellowtailF,Grenadier,Haddock,HakeS,HakeW,Halibut,Herring,Monkfish,Plaice,Pollock,Redfish,SharkB,SharkP,Skate,Turbot,Wolffish,Mackerel,SurfClam,Scallop,Squid,Whelk,Lobster,Shrimp,SnowCrab,JonahCrab,PMont,Swordfish,TunaBE,TunaBF,SharkM,Alewife,UNKFlounder,TunaAlb,TunaSJ,TunaYF,TunaUnk,SharkUnk),Argentine,'351',Cod,'100',Capelin,'360',Cusk,'173',Dogfish,'362',WinterF,'143',WitchF,'142',YellowtailF,'141',Grenadier,'179',Haddock,'110',HakeS,'172',HakeW,'171',Halibut,'130',Herring,'200',Monkfish,'177',Plaice,'140',Pollock,'170',Redfish,'120',SharkB,'372',SharkP,'369',Skate,'356',Turbot,'144',Wolffish,'174',Mackerel,'250',Swordfish,'251',TunaBE,'253',TunaBF,'254',SharkM,'375',Alewife,'350',SurfClam,'608',Scallop,'612',Squid,'613',Whelk,'617',Lobster,'700',Shrimp,'702',SnowCrab,'705',JonahCrab,'703',PMont,'712',UNKFlounder,'149',TunaAlb,'252',TunaSJ,'255',TunaYF,'256',TunaUnk,'259',SharkUnk,'379','000') main_species_caught, ",
  "to_number(substr(to_char(date_fished,'YYYY'),1,4)) ycaught, ",
  "to_number(substr(to_char(date_fished,'MM'),1,2)) mcaught, ",
  "to_number(substr(to_char(date_fished,'DD'),1,2)) dcaught, ",
  "0 trip_num,0 sub_trip_num,depth,decode((-1), ",
  "sign(depth-.01),'10', sign((depth-.01)*(depth-25.99)),'1', ",
  "sign((depth-25.995)*(depth-50.99)),'2', ",
  "sign((depth-50.995)*(depth-75.99)),'3', ",
  "sign((depth-75.995)*(depth-100.99)),'4', ",
  "sign((depth-100.995)*(depth-125.99)),'5', ",
  "sign((depth-125.995)*(depth-150.99)),'6', ",
  "sign((depth-150.995)*(depth-175.99)),'7', ",
  "sign((depth-175.995)*(depth-200.99)),'8', ",
  "sign((depth-200.995)*(depth-250.99)),'9', ",
  "sign((depth-250.995)*(depth-997.99)),'0','10') depthzone, ",
  "lat latitude,  lon longitude, 0 fish_days, hrs effort_hrs, ",
  "greatest(sets, tows, hooks) effort_count, 'XX' trip_sub_trip_flag, TotalWt, ",
  "Argentine,Capelin,Cod,Cusk,Dogfish,WinterF,WitchF,YellowtailF,Grenadier,Haddock,HakeS,HakeW,Halibut,Herring,Monkfish,Plaice,Pollock,Redfish,SharkB,SharkP,Skate,Turbot,Wolffish,Mackerel,SurfClam,Scallop,Squid,Whelk,Lobster,Shrimp,SnowCrab,JonahCrab,PMont,Swordfish,TunaBE,TunaBF,SharkM,Alewife,UNKFlounder,TunaAlb,TunaSJ,TunaYF,TunaUnk,SharkUnk ",
  "from cpueslip, cl.catchers_",Year, "  ",
  "where cfv = cfv_number (+)", sep=""))

#Reducing the species column to just the main species caught. If we want to model multiple species or investigate bycatch and species groups, we need to back up to the preceding 'long' tables.
#drop table cpuelogshort;
#create table cpuelogshort as
#select logcode,ekey,
#catchers_recid,region_code,cfv_number,yland,mland,dland,caught_date,nafo_division_code,nafo_unit_area,
#nvl(tonnage_class,0) tonnage_class, nvl(length_class,0) length_class,
#gear_type,main_species_caught,ycaught,mcaught,dcaught,
#trip_num,sub_trip_num,depthzone,latitude,longitude,
#fish_days,effort_hrs,effort_count,trip_sub_trip_flag,TotalWt,
#decode(main_species_caught,
#	'351',Argentine,'100',Cod,'360',Capelin,'173',Cusk,'362',Dogfish,
#	'143',WinterF,'142',WitchF,'141',YellowtailF,'179',Grenadier,
#	'110',Haddock,'172',HakeS,'171',HakeW,'130',Halibut,'200',Herring,'177',Monkfish,
#	'140',Plaice,'170',Pollock,'120',Redfish,'372',SharkB,'369',SharkP,
#	'356',Skate,'144',Turbot,'174',Wolffish,'250',Mackerel,'251',Swordfish,'350',Alewife,'608',SurfClam,'612',Scallop,'613',Squid,'617',Whelk,'700',Lobster,'702',Shrimp,'705',SnowCrab,'712',PMont,0) msc_tons
#from cpueloglong;
#drop table cpueslipshort;
#create table cpueslipshort as
#select logcode,ekey,
#catchers_recid,region_code,cfv_number,yland,mland,dland,caught_date,nafo_division_code,nafo_unit_area,
#nvl(tonnage_class,0) tonnage_class, nvl(length_class,0) length_class,
#gear_type,main_species_caught,ycaught,mcaught,dcaught,
#trip_num,sub_trip_num,depthzone,latitude,longitude,
#fish_days,effort_hrs,effort_count,trip_sub_trip_flag,
#(Argentine+Capelin+Cod+Cusk+Dogfish+WinterF+WitchF+YellowtailF+
#Grenadier+Haddock+HakeS+HakeW+Halibut+Herring+Monkfish+Plaice+Pollock+Redfish+SharkB+
#SharkP+Skate+Turbot+Wolffish+Mackerel+SurfClam+Scallop+Squid+Whelk+Lobster+Shrimp+SnowCrab+PMont+Swordfish+Alewife) TotalWt,
#decode(main_species_caught,
#	'351',Argentine,'100',Cod,'360',Capelin,'173',Cusk,'362',Dogfish,
#	'143',WinterF,'142',WitchF,'141',YellowtailF,'179',Grenadier,
#	'110',Haddock,'172',HakeS,'171',HakeW,'130',Halibut,'200',Herring,'177',Monkfish,
#	'140',Plaice,'170',Pollock,'120',Redfish,'372',SharkB,'369',SharkP,
#	'356',Skate,'144',Turbot,'174',Wolffish,'250',Mackerel,'251',Swordfish,'350',Alewife,'608',SurfClam,'612',Scallop,'613',Squid,'617',Whelk,'700',Lobster,'702',Shrimp,'705',SnowCrab,'712',PMont,0) msc_tons
#from cpuesliplong;

#Based on 2000-2002, the database requires 7 MB per year. The one-year table below includes log-only variables used by the Catch Rate application where slip data is problematic or missing.

sqlQuery(channel, paste("drop table cpue",Year,sep=""))
sqlQuery(channel, paste("create table cpue",Year," as ",
  "select l.catchers_recid,l.region_code,l.cfv_number,l.yland,l.mland,l.dland, ",
  "l.caught_date,l.nafo_division_code,l.nafo_unit_area,l.tonnage_class, l.length_class, ",
  "l.gear_type,s.main_species_caught,l.ycaught,l.mcaught,l.dcaught,l.trip_num, ",
  "l.sub_trip_num,l.depthzone,l.latitude,l.longitude,l.fish_days,l.effort_hrs, ",
  "l.effort_count,l.trip_sub_trip_flag,l.depth,s.TotalWt,l.Argentine,l.Capelin, ",
  "l.Cod,l.Cusk,l.Dogfish,l.WinterF,l.WitchF,l.YellowtailF,l.Grenadier,l.Haddock, ",
  "l.HakeS,l.HakeW,l.Halibut,l.Herring,l.Monkfish,l.Plaice,l.Pollock,l.Redfish,l.SharkB, ",
  "l.SharkP,l.Skate,l.Turbot,l.Wolffish,l.Mackerel,l.SurfClam,l.Scallop,l.Squid,l.Whelk, ",
  "l.Lobster,l.Shrimp,l.SnowCrab,l.JonahCrab,l.PMont,l.Swordfish,l.TunaBE,l.TunaBF, ",
  "l.SharkM,l.Alewife,l.UNKFlounder,l.TunaAlb,l.TunaSJ,l.TunaYF,l.TunaUnk,l.SharkUnk, ",
  "s.TotalWt sliptotwt,l.main_species_caught logmsc,l.TotalWt logtotwt ",
  "from cpueloglong l, cpuesliplong s ",
  "where l.logcode=s.logcode (+) and l.ekey=s.ekey (+) ",sep=""))

```

```{r}
sqlQuery(channel, "select * from agglog where rownum <5")
sqlQuery(channel, "select * from slip where rownum <5")
sqlQuery(channel, "select * from logadjslip where rownum <5")
sqlQuery(channel, "select * from cpuelog where rownum <5")
sqlQuery(channel, "select * from cpueslip where rownum <5")
sqlQuery(channel, "select * from cpueloglong where rownum <5")
sqlQuery(channel, "select * from cpuesliplong where rownum <5")
sqlQuery(channel, "select * from cpue1999 where rownum <5")
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
